Untwisting the Boerdijk--Coxeter Helix

### 2017-05-09 doc/TetrahelixUntwist.tex
improved proof by switching to radius

The vertices of the tetrahedra lie upon
three helices about the central axis.
The glussbot\cite{readglussbot} (or Tetrobot)\cite{TetrobotBook} uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot uses mechanical actuators which can change their length, connected by special joints,
such as the 3D printable Song-Kwon-Kim\cite{song2003spherical} joint, or the CMS joint\cite{HamlinSandersonCMS}
which allow many members to meet in a single point.
Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

@@

In principle in any three helices generated with \cref{eq:triple}
has at most nine distinct edge length classes. Each edge that connects
two rails potentially has a longer length and shorter length we donote with
a $+$ or $-$. So the classes are $\{ RR, BB, YY, RB_+, RB_-, BY_+, BY_-, RY_+, RY_- \}$.
If when projecting all vertices on the the $z$-axis, the interval
defined by the $z$ axis value of its endpoints contains no other vertices,
we call it a \emph{one-hop} edge, and if it does contain another vertex we
call it a \emph{two-hop} edge.
Then there are 
3 rail edges $\{ RR, BB, YY\}$,
3 one-hop lengths $\{ RB_-, BY_-, RY_- \}$ between each pair of 3 rails,
and 3 two-hop
lengths $\{ RB_{+}, BY_+, RY_+, \}$ between each pair of 3 rails,
where the two-hop length is at least the one-hop length.
However, if we generate the three helices symmetrically with \cref{eq:triple},
many of these lengths will be the same.
In fact, it is possible that there
will be only two distinct such classes.

@@

We seek a formula to generate optimal tetrahelices that accepts a
parameter that allows us to design the tetrahelix conveniently.
Please refer back to \cref{railanglefig}.
The pitch of the helix is an obvious choice, but is not defined when the
curvature is $0$, an important special case. The radius or the axial
distance between two nodes on the same rail are possible choices, but
perhaps the clearest choice is to build formulae that takes as their
input the ``rail angle'' $\rho$. We define $\rho$ to be the angle
formed in the X,Y plane $\angle H(0,0) O H(0,1)$ projecting out the $z$
axis and sighting along the positive $z$ axis. In other words, $\rho$
controls how far a rail edge of a tetrahelix deviates from being
parallel with the axis, or the ``twistiness'' of the tetrahelix. We use
the parameter $\chi = 1$ to indicate a chirality of counter-clockwise,
and $\chi = -1$ for clockwise.

@@

We would really like to know the partial derivative of the two-hop - one-hop with respect
to the radius to be able to understand how to choose the radius to form the minimimax optimum.

@@

Then:
\begin{align*}
  \text{two-hop} - \text{one-hop}  &= \sqrt{\frac{4}{9}  + r_{\rho}^2(g_{\rho}+ j_{\rho})}
  - \sqrt{\frac{1}{9} +r_{\rho}^2(f_{\rho}+k_{\rho}) }
\end{align*}

By graph inspection using Mathematica, we see the partial derivative of this with respect to
radius $r_{\rho}$ is always negative.
Since the partial derivative of $\text{two-hop} - \text{one-hop}$ with respect to the
radius $r_{\rho}$ is negative up until $\rho_{bc}$ where it is $0$, 
we opitmize the overall minimax distance by choosing the largest radius
up until one-hop $= 1$, the rail-edge length.

Therefore we decrease the minimax length
of the whole system as we increase the radius
up to the point that the shorter, one-hop distance is equal to the rail-length ($1$).
Therefore, to optimize the whole system so long as $\rho \leq \rho_{bc}$,
we equate one-hop to $1$ to find the optimum radius:

@@

Thus, by computing $r_{opt}$  and $d_{opt}$ as a function of $\rho$ from this equation, we can construct minimax optimal tetrahelix for an $0 \leq \rho \leq \rho_{bc}$.
\end{proof}

@@

If we look down the axis of an optimal tetrahelix as shown in \cref{axialview}, it happens that only one of the one-hop edges
(rendered in purple in our software)
comes closest to the axis. In other words, they define the radius of the incircle of the
projection, or the radius of a cylinder that would just fit inside the tetrahelix.
A formula for the inradius of the tetrahelix is useful if you are designing it as a structure that bears something internally,
such as a firehose, a pipe, or
a ladder for a human. The inradius $r_{in}(\rho)$ of
an optimal tetrahelix is a remarkably simple function of the radius $r$ and the rail angle $\rho$:
\begin{equation}
  \label{eq:inradius}
  r_{in}(\rho) = r \sin{\frac{\pi - \rho}{6}}
\end{equation}
Which can be seen from the trigonometry of a diagram of the projected one-hop edges
connecting four sequentially numbered vertices:

>>>>

The vertices of the tetrahedra lie upon
three helices about the central axis.
The \\
Glussbot\cite{readglussbot} (or Tetrobot)\cite{TetrobotBook} uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  These modular robot systems uses mechanical actuators which can change their length, connected by special joints,
such as the 3D printable Song-Kwon-Kim\cite{song2003spherical} joint in the case of the Glussbot, or the CMS joint\cite{HamlinSandersonCMS}, in the case of the Tetrobot,
which allow many members to meet in a single point.
Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

@@

In principle in any three helices generated with \cref{eq:triple}
has at most nine distinct edge length classes. Each edge that connects
two rails potentially has a longer length and shorter length we donote with
a $+$ or $-$. So the classes are $\{ RR, BB, YY, RB_+, RB_-, BY_+, BY_-, RY_+, RY_- \}$.
If when projecting all vertices onto (dropping the $x$ and $y$ coordinates) the $z$-axis, the interval
defined by the $z$ axis value of its endpoints contains no other vertices,
we call it a \emph{one-hop} edge, and if it does contain another vertex we
call it a \emph{two-hop} edge.
Then there are 
3 rail edges $\{ RR, BB, YY\}$,
3 one-hop lengths $\{ RB_-, BY_-, RY_- \}$ between each pair of 3 rails,
and 3 two-hop
lengths $\{ RB_{+}, BY_+, RY_+, \}$ between each pair of 3 rails,
where the two-hop length is at least the one-hop length.
However, if we generate the three helices symmetrically with \cref{eq:triple},
many of these lengths will be the same.
In fact, it is possible that there
will be only two distinct such classes.

@@

We seek a formula to generate optimal tetrahelices that accepts a
parameter that allows us to design the tetrahelix conveniently.
Please refer back to \cref{railanglefig}.
The pitch of the helix is an obvious choice, but is not defined when the
curvature is $0$, an important special case. The radius or the axial
distance between two nodes on the same rail are possible choices, but
perhaps the clearest choice is to build formulae that takes as their
input the ``rail angle'' $\rho$. We define $\rho$ to be the angle
formed in the X,Y plane $\angle H(0,0) O H(0,1)$ projecting out the $z$
axis and sighting along the positive $z$ axis. In other words, $\rho$
controls how far a rail edge of a tetrahelix deviates from being
parallel with the axis, or the ``twistiness'' of the tetrahelix. We use
the parameter $\chi = 1$ to indicate a chirality of counter-clockwise,
and $\chi = -1$ for clockwise.
%% I have to add here whether I am using a left-handed or right-handed coordinate
%% system.

@@

By definition of minimax edge length optimality, we are trying to minimize:
\[
\frac{\max{\{1,\text{one-hop}(r),\text{two-hop}(r)}\}}{\min{\{1,\text{one-hop}(r),\text{two-hop}(r)}\}}
\]
But since $\text{two-hop}(r) >= \text{one-hop}(r)$, this is equivalent to:
\[
 \frac{   \max \{1,\text{two-hop}(r)\}}
      {   \min \{1,\text{one-hop}(r)\}}
\]
This quantity will be equal to one of:
\[ \frac{\text{two-hop}(r)}{1},
\frac{1}{\text{one-hop}(r)},
\frac{\text{two-hop}(r)}{\text{one-hop}(r)}
\]

We know that both one-hop($r$) and two-hop($r$) increase monotonically and continuously with increasing $r$.
By inspection it seems likely that we will minize this set by equating $\text{one-hop}(r)$ or $\text{two-hop}(r)$
to 1, but to be absolutely sure we must examine the partial derivative of the ratio
 $\frac{\text{two-hop}(r)}{\text{one-hop}(r)}$ in this range.

Although complicated, we can use Mathematica to investigate
the partial derivative of the $\text{two-hop}(r) - \text{one-hop}(r)$ with respect
to the radius to be able to understand how to choose the radius to form the minimax optimum.

@@

Then:
\begin{align*}
  \frac{\text{two-hop}}{ \text{one-hop}}  &=
  \frac{\sqrt{\frac{4}{9}  + r_{\rho}^2(g_{\rho}+ j_{\rho})}}
       {\sqrt{\frac{1}{9} +r_{\rho}^2(f_{\rho}+k_{\rho}) }}
\end{align*}

By graph inspection using Mathematica, we see the partial derivative of this with respect to
radius $r_{\rho}$ is always negative.
Since the partial derivative of $\text{two-hop} / \text{one-hop}$ with respect to the
radius $r_{\rho}$ is negative up until $\rho_{bc}$, this ratio goes down
as the radius goes up, and 
we optimize the overall minimax distance by choosing the largest radius
up until one-hop $= 1$, the rail-edge length.

Therefore we decrease the minimax length
of the whole system as we increase the radius
up to the point that the shorter, one-hop distance is equal to the rail-length, $1$.
Therefore, to optimize the whole system so long as $\rho \leq \rho_{bc}$,
we equate one-hop to $1$ to find the optimum radius:

@@

Thus, by computing $r_{opt}$  and $d_{opt}$ as a function of $\rho$ from this equation, we can construct minimax optimal tetrahelix for an $0 \leq \rho \leq \rho_{bc}$.
%% Add here references to these equations.
\end{proof}

@@

If we look down the axis of an optimal tetrahelix as shown in \cref{axialview}, it happens that only
the one-hop edges
(rendered in purple in our software)
comes closest to the axis. In other words, they define the radius of the incircle of the
projection, or the radius of a cylinder that would just fit inside the tetrahelix.
A formula for the inradius of the tetrahelix is useful if you are designing it as a structure that bears something internally,
such as a firehose, a pipe, or
a ladder for a human. The inradius $r_{in}(\rho)$ of
an optimal tetrahelix is a remarkably simple function of the radius $r$ and the rail angle $\rho$:
\begin{equation}
  \label{eq:inradius}
  r_{in}(\rho) = r \sin{\frac{\pi - \rho}{6}}
\end{equation}
Which can be seen from the trigonometry of a diagram of the projected one-hop edges
connecting four sequentially numbered vertices:
